---
title: "Analysis"
output: html_document
date: "2023-12-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load library

```{r}
library(tidyverse)
library(lme4)
library(ggbeeswarm)
library(brms)
options(brms.backend = "cmdstanr")
```

## Read in participant data

```{r}
df.subj.aug <- read_csv("../data/participants_aug2023.csv")
df.subj.dec <- read_csv("../data/participants_dec2023.csv")

df.subj <- bind_rows(df.subj.aug, df.subj.dec) %>%
  group_by(sonaID) %>%
  summarize(across(everything(), last))  # only keep the lastest entry

head(df.subj)
```


## Read in duration data

```{r}
df.dur <- read_csv("../data/audio_duration_CVenvironment.csv") %>%
  mutate(Filename = str_sub(Filename, end = -5))
df.dur
```

## Read in experiment data
- Trial exclusion criteria:
  - Trials without responses
  - Trials whose RT is shorter than stimulus duration
  - Trials whose log(RT) is >= 2 standard devations within each participant+talker group

- Participant exclusion criteria:
  - Participants with SLH == yes
  - Participants who don't speak English natively
  - Participants whose AoA for English is > 5
  - Participants whose valid trials are < 75% of total trials

```{r}
df.raw <- read_csv("../data/fric.categorization.df.csv") %>%
  left_join(x = ., y = df.subj, by = c("id" = "sonaID")) %>%
  
  # only include participants whose English is native and age of acquisition is <= 5
  filter(SLH == 'No', EnglishNative == 'Yes', EN_AOA <= 5) %>%
  dplyr::select(-word, -CategorizeWhat) %>%
  mutate(filename = str_sub(filename, end = -5)) %>%
  left_join(x = ., y = df.dur, by = c("filename" = "Filename")) %>%
  mutate(
    TYPE = if_else(CategoryLabel %in% c("S", "SH"), "C", "CV"),
    CONS = str_extract(CategoryLabel, "(SH|S)"),
    CORRECT = Response == CONS,
    
    # trials without response will not be kept
    KEPT = if_else(is.na(Response), 0, 1),
    Duration = Duration * 1000,
    
    # trials whose RT is shorter than the stimulus duration will not be kept
    KEPT = if_else(KEPT == 1, if_else(Duration <= RT, 1, 0), 0),
    LOG_RT = if_else(KEPT == 1, log(RT), NA)
  ) %>%
  dplyr::select(-CategoryLabel) %>%
  group_by(id, talker) %>%
  mutate(
    SE = abs((LOG_RT - mean(LOG_RT, na.rm = TRUE)) / sd(LOG_RT, na.rm = TRUE)),
    
    # trials with standard error >= 2 will not be kept
    KEPT = if_else(KEPT == 1, if_else(SE >= 2, 0, 1), 0),
    
    # removed participants of whom more than 75% of trials are not valid
    KEPT = if_else(id %in% c(38371, 38689, 38677, 38644, 38038, 37966, 35083, 37948, 38686, 36430), 0, KEPT)
  )

head(df.raw)

df.raw %>% write_csv(file = "../data/raw.csv")
df.raw <- read_csv("../data/raw.csv")
```

## Visualization

```{r}
# by-talker production distance
# mean (SD)
# 159: 13.07 (1.99)
# 2470: 11.63 (1.16)
# 115: 10.58 (1.06)
# 2471: 2.65 (0.38)
# 2453: 2.49 (0.36)
# 2436: 1.99 (0.31)

df.raw %>%
  filter(KEPT == 1) %>%
  mutate(talker = as.factor(talker),
         talker = fct_relevel(talker, "159", "2470", "115", "2471", "2453", "2436")) %>%
  group_by(talker, TYPE, id, CONS) %>%
  summarize(MEAN_CORRECT = mean(CORRECT)) %>%
  ggplot(aes(x = as.factor(talker), y = MEAN_CORRECT)) +
  theme_bw() +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.5) +
  facet_grid(CONS ~ TYPE)
```

```{r}
df.raw %>%
  filter(KEPT == 1) %>%
  mutate(talker = as.factor(talker),
         talker = fct_relevel(talker, "159", "2470", "115", "2471", "2453", "2436")) %>%
  ggplot(aes(x = as.factor(talker), y = RT)) +
  theme_bw() +
  geom_violin() +
  facet_grid(CONS ~ TYPE)
```

## Stats analysis

```{r}
df.full <- df.raw %>%
  filter(KEPT == 1) %>%
  mutate(
    talker = as.factor(talker),
    
    # using 2471 as the ref level because this talker has the lowest accuracy
    talker = fct_relevel(talker, "2471")
  )

# TODO: look into divergent transitions with more complex models
# use a very simple random structure here because more complex ones led to divergent transitions
# use default priors for now
# m_acc <- brm(formula = CORRECT ~ TYPE * CONS * talker + (1 | id) + (1 | word), data = df.full,
#              family = bernoulli(link = "logit"), chains = 4, cores = 4)
# 
# saveRDS(m_acc, "../model/m.acc.random.intercept.rds")
m_acc <- readRDS("../model/m.acc.random.intercept.rds")
summary(m_acc)
```

```{r}
# m_RT <- brm(formula = LOG_RT ~ TYPE * CONS * talker + (1 | id) + (1 | word), data = df.full,
#             chains = 4, cores = 4)
# 
# saveRDS(m_RT, "../model/m.rt.random.intercept.rds")
m_RT <- readRDS("../model/m.rt.random.intercept.rds")
summary(m_RT)
```